---
title: "Additional Analysis: Risk-Stratified Screening and Re-invitation Strategies"
author: "Mary Ann Binuya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document: default
---

# Aims
1. To describe observed detection patterns under the current biennial screening strategy.
2. To evaluate how detection mode varies by predicted 5-year BOADICEA risk using 4 model configurations.
3. To simulate risk-based re-invitation strategies and assess potential clinical utility compared to fixed-interval screening.

```{r setup, include=FALSE}
#Set libpath
#.libPaths(c('U:/libraries', .libPaths()))

knitr::opts_chunk$set(
  echo = TRUE,
  fig.retina = 3,
  fig.path = "Output/6 Simulations/",
  warning = FALSE)

library(dplyr) #for data manipulations
library(table1) #for table1
library(survival) #for survival analysis
library(survminer) #for plot
library(ggplot2) #for plotting
library(purrr) #for better plots

rm(list = ls())
```

# Load required datasets
```{r load}
load("WS_03_BOADICEA_output.RData")

   dat <- dat_full %>%
    mutate(predrisk_mod1 = BOA_5yrsrisk_Model1/100,
            predrisk_mod2 = BOA_5yrsrisk_Model2/100,
            predrisk_mod3a = BOA_5yrsrisk_Model3a/100,
            predrisk_mod3b = BOA_5yrsrisk_Model3b/100,
            predrisk_mod4 = BOA_5yrsrisk_Model4/100,
            predrisk_mod5a = BOA_5yrsrisk_Model5a/100,
            predrisk_mod5b = BOA_5yrsrisk_Model5b/100) %>%
     filter(!(bc == 1 & omgedrag == "In situ")) #remove all DCIS from cases, n=1089
   
  rm(list=setdiff(ls(), c("dat")))
```

# 1. Describe observed detection patterns under the current biennial screening strategy.
```{r det_mode}
# Select cases only
cases <- subset(dat, bc == 1)

# Apply labels per NKR codebook
cases$det_mode <- factor(cases$scrrel,
  levels = c(1, 2, 3, 4, 5),
  labels = c("Screen-detected\n(≤12 months, with recall)",
    "Interval cancer\n(≤30 months, no recall)",
    "Pseudo-interval cancer\n(>30 months, no recall)",
    "Unclear\n(recall but wrong BI-RADS or side)",
    "Unclear\n(recall but diagnosis >12 months)"))

res_detmode <- cases %>% group_by(det_mode) %>%
  summarise(Count = n()) %>%
  mutate(Proportion = round(Count / sum(Count) * 100, 1))
res_detmode
# Same for imputed and unimputed data
```

# 2. Evaluate how detection mode varies by predicted 5-year BOADICEA risk using 4 input configurations.
## 2.1 Pseudo and interval cancers separated
```{r det_mode_byboa}
# Drop unclear categories:
# 3 women (Unclear (recall but wrong BI-RADS or side)) + 1 woman (Unclear (recall but diagnosis >12 months))
cases_clean <- cases %>% filter(scrrel %in% c(1, 2, 3)) # 314 cases left

# Define function to categorize risk
categorize_risk <- function(risk) {
  cut(risk,
      breaks = c(-Inf, 0.0167, 0.03, Inf),
      labels = c("Low\n(<1.67%)", "Intermediate\n(1.67-3%)", "High\n(>3%)"),
      right = FALSE)}

# List of BOADICEA model to be analyzed
boadicea_models <- c("predrisk_mod2", "predrisk_mod3b", "predrisk_mod4", "predrisk_mod5b")
names(boadicea_models) <- c("Model 2: Age + Clinical",
                            "Model 3b: Age + Clinical + BD",
                            "Model 4: Age + Clinical + PRS",
                            "Model 5b: Age + Clinical + PRS + BD")

# Results
res_detmode_BOA <- list()

for (model in names(boadicea_models)) {
  risk_var <- boadicea_models[model]
  
  temp <- cases_clean %>%
    mutate(risk_group = categorize_risk(.data[[risk_var]])) %>%
    group_by(risk_group, det_mode) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(risk_group) %>%
    mutate(Proportion = round(n / sum(n) * 100, 1)) %>%
    mutate(Model = model)
  
  res_detmode_BOA[[model]] <- temp
}

# Results
res_det_BOA <- bind_rows(res_detmode_BOA) %>%
  select(Model, risk_group, det_mode, n, Proportion)

# Label
res_det_BOA$risk_group <- factor(res_det_BOA$risk_group,
                                            levels = c("Low\n(<1.67%)", "Intermediate\n(1.67-3%)", "High\n(>3%)"))


# Plot: Stacked bar by risk group and detection mode

ggplot(res_det_BOA, aes(x = risk_group, y = Proportion, fill = det_mode)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  facet_wrap(~ Model, ncol = 2) +
  scale_fill_brewer(palette = "Set2", name = "Detection Mode") +
  labs(
    title = "Detection Mode Distribution by Risk Group and BOADICEA Model",
    x = "BOADICEA Risk Group",
    y = "Proportion (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5)
  )

  rm(temp, risk_var, res_detmode_BOA)
```

## 2.2 Pseudo and interval cancers combined as non-screen-detected cancers
```{r det_mode_byboa_combined}
# Drop unclear categories:
# 3 women (Unclear (recall but wrong BI-RADS or side)) + 1 woman (Unclear (recall but diagnosis >12 months))
cases_clean <- cases %>%
  filter(scrrel %in% c(1, 2, 3)) %>%  # Exclude unclear cases
  mutate(det_mode2 = case_when(
    scrrel == 1 ~ "Screen-detected\n(≤12 months)",
    scrrel %in% c(2, 3) ~ "Interval cancers\n(>12 months and no recall)"
  ))

# Results
res_detmode_BOA <- list()

for (model in names(boadicea_models)) {
  risk_var <- boadicea_models[model]
  
  temp <- cases_clean %>%
    mutate(risk_group = categorize_risk(.data[[risk_var]])) %>%
    group_by(risk_group, det_mode2) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(risk_group) %>%
    mutate(Proportion = round(n / sum(n) * 100, 1)) %>%
    mutate(Model = model)
  
  res_detmode_BOA[[model]] <- temp
}

# Results
res_det_BOA2 <- bind_rows(res_detmode_BOA) %>%
  select(Model, risk_group, det_mode2, n, Proportion)

# Label
res_det_BOA2$det_mode2 <- factor(res_det_BOA2$det_mode2,
                                          levels = c("Screen-detected\n(≤12 months)",
                                                     "Interval cancers\n(>12 months and no recall)"))

res_det_BOA2$risk_group <- factor(res_det_BOA2$risk_group,
                                            levels = c("Low\n(<1.67%)", "Intermediate\n(1.67-3%)", "High\n(>3%)"))


# Plot: Stacked bar by risk group and detection mode

ggplot(res_det_BOA2, aes(x = risk_group, y = Proportion, fill = det_mode2)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  facet_wrap(~ Model, ncol = 2) +
  scale_fill_brewer(palette = "Set2", name = "Detection Mode") +
  labs(
    title = "Detection Mode Distribution by Risk Group and BOADICEA Model",
    x = "BOADICEA Risk Group",
    y = "Proportion (%)"
  ) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom",
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0.5)
  )

  rm(temp, risk_var, res_detmode_BOA)
  rm(boadicea_models, model, categorize_risk)
```

# 3. Simulate risk-based re-invitation strategies and assess potential clinical utility compared to fixed-interval screening.
## 3.1 What happens to detection efficiency and burden under different screening frequencies?
```{r simulate_screens}
# Prepare dataset
tp <- 5  # time horizon
dat_screen <- dat  # copy for truncation
dat_screen$bc_trunc <- dat_screen$bc  # initialize with original bc
dat_screen$bc_trunc[dat_screen$futime > tp] <- 0 # Set bc = 0 if futime > 5
dat_screen$futime_trunc <- pmin(dat_screen$futime, tp) # Truncate futime at 5 years
dat_screen$wt <- dat_screen$wt_KM

# Sanity check
# dat_screen %>% filter(futime>5 & bc==1)
# summary(dat_screen$futime_trunc)
# table(dat_screen$bc, dat_screen$bc_trunc, useNA="always")

# Define fixed screening intervals within 5 years
screen_schedules <- list(
  Annual = c(0, 1, 2, 3, 4),
  Biennial = c(0, 2, 4),
  Triennial = c(0, 3)
)

# Function: screen-detected if diagnosis within 1 year of any screening round
is_screen_detected <- function(ftime, screen_times, window) {
  any(sapply(screen_times, function(t) ftime >= t & ftime < t + window))
}

# Function: simulate detection mode under a screen strategy
simulate_detection_timing <- function(data, strategy_name, screen_times) {
  window <- if (strategy_name == "Annual") 0.75 else 1 #to prevent overlap between successive screens
  
  data_sim <- data %>%
    mutate(
      detection_mode_sim = case_when(
        bc_trunc == 1 & sapply(futime_trunc, is_screen_detected, screen_times = screen_times, window = window) ~ "Screen-detected",
        bc_trunc == 1 ~ "Interval cancer",
        bc_trunc == 0 ~ "No cancer"
      )
    )

  # Weighted summaries
  N_cancers <- sum(data_sim$wt[data_sim$bc_trunc == 1])
  N_no_cancers <- sum(data_sim$wt[data_sim$detection_mode_sim == "No cancer"])
  N_screen_detected <- sum(data_sim$wt[data_sim$detection_mode_sim == "Screen-detected"])
  N_interval <- sum(data_sim$wt[data_sim$detection_mode_sim == "Interval cancer"])

  total_screens <- sum(sapply(screen_schedules[[strategy_name]], function(t) {
    sum(data_sim$futime_trunc >= t)})) #count only screens where woman was alive and at risk
  interval_rate <- round(N_interval / N_cancers * 100, 1)
  screens_per_cancer <- round(total_screens / N_screen_detected, 1)

  data.frame(
    Strategy = strategy_name,
    Total_Cancers = round(N_cancers),
    Total_No_Cancers = round(N_no_cancers),
    Screen_Detected = round(N_screen_detected),
    Interval_Cancers = round(N_interval),
    Interval_Cancer_Rate = interval_rate,
    Total_Screens = total_screens,
    Screens_per_Cancer = screens_per_cancer
  )
}

# Apply simulation once (point estimates)
results_simulation <- lapply(names(screen_schedules), function(name) {
  simulate_detection_timing(dat_screen, name, screen_schedules[[name]])
}) %>% bind_rows()

# Bootstrap for uncertainty (CIs)
set.seed(123)
n_boot <- 200
boot_results <- list()

for (b in 1:n_boot) {
  boot_data <- dat_screen[sample(nrow(dat_screen), replace = TRUE), ]
  res_b <- lapply(names(screen_schedules), function(name) {
    simulate_detection_timing(boot_data, name, screen_schedules[[name]])
  }) %>% bind_rows() %>% mutate(bootstrap = b)
  boot_results[[b]] <- res_b
}

boot_df <- bind_rows(boot_results)

# Summarize key outcomes with CIs
summary_stats <- boot_df %>%
  group_by(Strategy) %>%
  summarise(
    Total_Screens_median = median(Total_Screens),
    Total_Cancers_median = median(Total_Cancers),
    Total_No_Cancers_median = median(Total_No_Cancers),
    
    Screen_Detected_median = median(Screen_Detected),
    Screen_Detected_lower = quantile(Screen_Detected, 0.025),
    Screen_Detected_upper = quantile(Screen_Detected, 0.975),

    Interval_Cancers_median = median(Interval_Cancers),
    Interval_Cancers_lower = quantile(Interval_Cancers, 0.025),
    Interval_Cancers_upper = quantile(Interval_Cancers, 0.975),

    Interval_Cancer_Rate_median = median(Interval_Cancer_Rate),
    Interval_Cancer_Rate_lower = quantile(Interval_Cancer_Rate, 0.025),
    Interval_Cancer_Rate_upper = quantile(Interval_Cancer_Rate, 0.975),

    Screens_per_Cancer_median = median(Screens_per_Cancer),
    Screens_per_Cancer_lower = quantile(Screens_per_Cancer, 0.025),
    Screens_per_Cancer_upper = quantile(Screens_per_Cancer, 0.975)) %>%
  ungroup()

# Final table format
res_simulation <- summary_stats %>%
  transmute(
    Strategy,
    `Total Screens` = round(Total_Screens_median, 0),
    `Total Cancers` = round(Total_Cancers_median, 0),
    `No Cancers` = round(Total_No_Cancers_median, 0),
    `Screen-Detected` = sprintf("%.0f (%.0f–%.0f)",
                                Screen_Detected_median, Screen_Detected_lower, Screen_Detected_upper),
    `Interval Cancers` = sprintf("%.0f (%.0f–%.0f)",
                                 Interval_Cancers_median, Interval_Cancers_lower, Interval_Cancers_upper),
    `Interval Cancer Rate (%)` = sprintf("%.1f (%.1f–%.1f)",
                                         Interval_Cancer_Rate_median, Interval_Cancer_Rate_lower, Interval_Cancer_Rate_upper),
    `Screens per Cancer` = sprintf("%.1f (%.1f–%.1f)",
                                   Screens_per_Cancer_median, Screens_per_Cancer_lower, Screens_per_Cancer_upper))

res_simulation

rm(list = setdiff(ls(), c("dat", "cases", "case_clean", "res_detmode", "res_det_BOA", "res_det_BOA2",
                          "dat_screen", "tp", "res_simulation")))
```

## 3.2 Does using predicted risk to decide who to invite result in better overall outcomes than treating all or none?
```{r simulate_NB}
### Marginal strategies – risk-informed one-time re-invitations

# Evaluate Net Benefit (NB) of assigning a one-time re-invitation based on baseline 5-year predicted risk.
# High-risk women (>3%) are re-invited at year 1, low-risk women (<1.67%) at year 3.
# Comparator: screen-all at year 2.
# All women receive a baseline screen at year 0. Outcomes are evaluated within a fixed detection window.

# Prepare data
dat_screen <- dat %>%
  mutate(
    futime_trunc = pmin(futime, 5),
    bc_trunc = ifelse(futime > 5, 0, bc),
    wt = wt_KM
  )

# Define thresholds
low_thresh <- 0.0167
high_thresh <- 0.03
inter_thresh <- (low_thresh + high_thresh)/2 #0.02335

# Function: Risk-informed one-time re-invitation
compute_NB_stratified <- function(data, risk_score,
                                  strategy = c("low-risk", "high-risk"),
                                  low_thresh, high_thresh) {
  strategy <- match.arg(strategy)

  data <- data %>%
    mutate(
      risk_value = .data[[risk_score]],
      eligible = case_when(
        strategy == "low-risk" ~ risk_value < low_thresh,
        strategy == "high-risk" ~ risk_value > high_thresh,
        TRUE ~ FALSE
      )
    ) %>%
    filter(eligible)

  # Assign screening year and detection window
  if (strategy == "high-risk") {
    screen_year <- 1
    window <- 0.75
    pt <- high_thresh
  } else if (strategy == "low-risk") {
    screen_year <- 3
    window <- 1.0
    pt <- low_thresh
  }

  # Apply detection logic
  data <- data %>%
    rowwise() %>%
    mutate(
      det_time = if (bc_trunc == 1) futime_trunc else NA_real_,
      screen_detected = ifelse(
        bc_trunc == 1 &
        futime_trunc >= screen_year & futime_trunc < screen_year + window,
        1, 0
      ),
      false_pos = ifelse(
        bc_trunc == 0 &
        futime_trunc >= screen_year & futime_trunc < screen_year + window,
        1, 0
      )
    ) %>%
    ungroup() %>%
    filter(!(bc_trunc == 0 & futime_trunc < screen_year))  # Exclude controls not alive at screening

  TP <- sum(data$wt * data$screen_detected, na.rm = TRUE)
  FP <- sum(data$wt * data$false_pos, na.rm = TRUE)
  N <- sum(data$wt, na.rm = TRUE)

  NB <- TP / N - FP / N * (pt / (1 - pt))

  data.frame(
    Strategy = strategy,
    TP = round(TP),
    FP = round(FP),
    NB_model = round(NB * 1000, 1)
  )
}

# Function: Screen-all at year 2 comparator
compute_NB_screen_all <- function(data, risk_score, pt = inter_thresh, screen_year = 2, window = 1.0) {
  data <- data %>%
    rowwise() %>%
    mutate(
      det_time = if (bc_trunc == 1) futime_trunc else NA_real_,
      screen_detected = ifelse(
        bc_trunc == 1 &
        futime_trunc >= screen_year & futime_trunc < screen_year + window,
        1, 0
      ),
      false_pos = ifelse(
        bc_trunc == 0 &
        futime_trunc >= screen_year & futime_trunc < screen_year + window,
        1, 0
      )
    ) %>%
    ungroup() %>%
    filter(!(bc_trunc == 0 & futime_trunc < screen_year))  # Exclude controls not alive at screening

  TP <- sum(data$wt * data$screen_detected, na.rm = TRUE)
  FP <- sum(data$wt * data$false_pos, na.rm = TRUE)
  N <- sum(data$wt, na.rm = TRUE)

  NB <- TP / N - FP / N * (pt / (1 - pt))

  data.frame(
    Strategy = "Screen all (year 2)",
    TP = round(TP),
    FP = round(FP),
    NB_model = round(NB * 1000, 1)
  )
}

# Model inputs
model_cols <- c("predrisk_mod2", "predrisk_mod3b", "predrisk_mod4", "predrisk_mod5b")
model_names <- c("Model 2: Age + Clinical",
                 "Model 3b: Age + Clinical + BD",
                 "Model 4: Age + Clinical + PRS",
                 "Model 5b: Age + Clinical + PRS + BD")

# Simulate all strategies for each model
nb_stratified_results <- map2_dfr(model_cols, model_names, function(risk_col, model_name) {
  bind_rows(
    compute_NB_stratified(dat_screen, risk_col, "low-risk", low_thresh, high_thresh),
    compute_NB_stratified(dat_screen, risk_col, "high-risk", low_thresh, high_thresh)
  ) %>% mutate(Model = model_name)
})

# Simulate screen-all comparator
nb_screen_all_results <- map2_dfr(model_cols, model_names, function(risk_col, model_name) {
  compute_NB_screen_all(dat_screen, risk_col) %>%
    mutate(Model = model_name)
})

# Combine all results
nb_all_results <- bind_rows(nb_stratified_results, nb_screen_all_results) %>%
  mutate(NB_ScreenNone = 0) %>%
  select(Model, Strategy, TP, FP, NB_model, NB_ScreenNone)

# View results
nb_all_results
```

## 3.2 Scale the NB results
```{r nb_prop}
# Calculate proportions of high-risk and low-risk women per model
risk_group_props <- map2_dfr(model_cols, model_names, function(risk_col, model_name) {
  dat_screen %>%
    mutate(risk_value = .data[[risk_col]]) %>%
    summarise(
      Proportion_high = mean(risk_value > high_thresh, na.rm = TRUE),
      Proportion_low  = mean(risk_value < low_thresh, na.rm = TRUE)
    ) %>%
    mutate(Model = model_name)
})

risk_group_props

# Step 2: Merge proportions with NB results
nb_standardized <- nb_all_results %>%
  left_join(risk_group_props, by = "Model") %>%
  mutate(
    Proportion = case_when(
      Strategy == "high-risk" ~ round(Proportion_high, 2),
      Strategy == "low-risk" ~ round(Proportion_low, 2),
      Strategy == "Screen all (year 2)" ~ 1.0,
      TRUE ~ NA_real_
    ),
    Standardized_NB = round(NB_model * Proportion, 1)
  ) %>%
  select(Model, Strategy, TP, FP, NB_model, Proportion, Standardized_NB)

nb_standardized
```


```{r save}
save.image("WS_06_Simulations.RData")
```
